Clustering Constraints on the Relative Sizes of Central and Satellite Galaxies

### 2017-11-08 paper/galsize_paper.tex
edited until S 5.2

Inclusion of satellite mass loss in the model introduces correlations in the scatter: after mass loss, satellites have smaller sizes than centrals of the same $\mstar$ due to post-infall stripping.
The purple curves in Figure \ref{fig:mstarmodelclustering} show that satellite mass loss impacts $\rhalf-$dependent clustering in the expected manner. The ``small" galaxy population has a higher satellite fraction due to satellites being smaller than centrals of the same $\mstar,$ and so in this model small galaxies cluster more strongly relative to large. However, the magnitude of the effect is not sufficiently strong  to produce clustering predictions that are consistent with observations. Evidently it is difficult to strip enough mass from satellites so that the $\rhalf-$dependent clustering predictions are in agreement with observations. As we argue in \S\ref{sec:interpretation}, this supports the conclusion that the relative difference between central and satellite size is largely in place at the time of satellite infall.

@@

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/rvir_only_lensing_prediction.pdf}
\caption{
{\bf Prediction for $\rhalf-$dependence of galaxy lensing.}
Using the $\rvir-$based model described in \S\ref{subsubsec:rvirmodel}, we make predictions for as-yet-unseen measurements of the $\rhalf-$dependence of galaxy lensing of $\mstar-$complete samples, where ``small" and ``large" subsamples are defined as in \S\ref{subsec:sizedef}. As a generic consequence of satellites being smaller than centrals of the same mass, we predict that future lensing measurements of $\mstar-$complete samples will show $\Delta\Sigma$ of small galaxies to be significantly stronger relative to large galaxies of the same stellar mass. To date, the $\rhalf-$dependence of $\Delta\Sigma$ has only been measured for color-selected samples \citep{charlton_etal17}, in which the trend is reversed and weakened: for both blue and red samples, $\Delta\Sigma$ of large galaxies is (slightly) stronger relative to small galaxies. 
}
\label{fig:lensingprediction}
\end{figure}
%-----------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/rvir_only_cen_sat_sizes.pdf}
\caption{
{\bf Relative sizes of centrals and satellites.}
In a narrow bin of halo mass $\mhalo\equiv\mpeak\approx10^{12}\msun,$ we show the distribution of model galaxy sizes for different subpopulations galaxies, as predicted by the $\rvir-$based model. The red histogram shows the sizes of satellites; the blue histogram shows host halos that have never passed inside the virial radius of a larger halo (``true centrals"); the green histogram host halos that were subhalos inside a larger at some point in their past history (``splashback centrals"). In the $\rvir-$based model, galaxy size is set by the {\em physical} size of the virial radius at the time the halo attains its peak mass, naturally resulting in smaller sizes for satellites and splashback centrals relative to true centrals of the same $\mhalo.$
}
\label{fig:censatsizehist}
\end{figure}
%-----------------------------------------------------------------------------------------------------

\section{A Unified Interpretation of $\rhalf-$Dependent Clustering}
\label{sec:interpretation}

The clustering measurements shown in the previous section present (at least) two intriguing puzzles. First, the modeling assumption $\rhalf\propto\rvir$ predicts size-dependent galaxy clustering that is in quite good agreement with observations, particularly considering the simplicity of the model. Yet, small galaxies cluster more strongly relative to large, which seems at odds with naive expectations based on halo mass, since $\rvir\propto\mvir.$ Second, there appears to be a highly non-trivial dependence of $\rhalf-$dependent clustering upon whether or not the galaxy sample has been selected by color. We propose a simple and unified solution to both puzzles in \S\ref{subsec:censat_sizes}, and discuss the broader implications of our interpretation in \S\ref{subsec:broader_implications}.

\subsection{The Relative Size of Centrals vs. Satellites}
\label{subsec:censat_sizes}

A straightforward resolution to the first puzzle is shown in Figure \ref{fig:censatsizehist}, which compares the $\rhalf$ distributions of central, satellite, and splashback galaxies with the same halo mass $\mhalo\equiv\mpeak\approx10^{12}\msun.$ A ``splashback central"  is defined as a present-day central that used to be a satellite, i.e., its main progenitor halo passed inside the virial radius of a larger halo at some point in its past history. On the other hand, we define a ``true central" as a galaxy that has never been a satellite.

@@

We can also understand the $\mstar-$dependence of Figure \ref{fig:clustering_ratio_upshot} in term of the relative sizes of centrals vs. satellites. The satellite fraction $F_{\rm sat}(\mstar)$ decreases as $\mstar$ increases \citep[e.g.,][]{guo_etal11,reddick_etal13}. As $F_{\rm sat}$ decreases, there are fewer satellites available to preferentially weight the subsample of smaller galaxies. Thus models where $\rhalf-$dependent clustering is driven by satellite vs. central size naturally predict the observed $\mstar-$dependence of the signal, without introducing any explicitly $\mstar-$dependent ingredients in the size model.

Upon first glance, the $\rhalf-$dependence of color-selected samples shown in Figure \ref{fig:colorclustering} appear to be driven by different principles, since the trends substantially weaken and change sign. However, even very simple models based entirely on central vs. satellite size differences can give rise to this behavior. For example, it is well known that satellites are redder than centrals of the same stellar mass \citep[e.g.,][]{vdB_etal08}; accordingly, red galaxies cluster much stronger relative to blue \citep{zehavi_etal11}. Pre-selecting galaxy samples by color will thus significantly weaken the dynamic range of $F_{\rm sat}$ within each color-selected subsample. This reduction in the role of satellites could be sufficient for {\em central} galaxies to be responsible for the dominant effect. When the influence of centrals dominates the signal, models where $\rhalf\propto\rvir$ predict that $\rhalf-$dependent clustering should reverse sign, since $\rvir\propto\mhalo^{1/3},$ and halo bias $b(\mhalo)$ increases with halo mass. Models where $\rhalf\propto\rvir$ would also predict this reversed signal should have relatively small magnitude, since $\rvir$ scales only weakly with $\mhalo,$ and because $b(\mhalo)$ is weak function of mass for halos hosting galaxies $\mstar\lesssim10^{11}\msun.$ Therefore, if the segregation of galaxies into red- and blue-sequences is correlated sufficiently strongly with central vs. satellite designation, then the $\rhalf-$dependent clustering of color-selected samples shown in the top panels of Figure \ref{fig:colorclustering} are consistent with predictions of models where $\rhalf\propto\rvir.$

While we consider such an explanation plausible, a conclusive determination of the principles underlying the clustering trends shown in Figure \ref{fig:colorclustering} requires proper forward modeling of size and color jointly. Such an effort is beyond the scope, but well-motivated, by the present work. See Section \ref{subsec:broader_implications} for further discussion.

@@

The clustering ratios in Figure \ref{fig:mstarmodelclustering} show that the impact of physically motivated levels of satellite mass loss likely cannot explain the $\rhalf-$dependent clustering seen in SDSS. As an additional test, we reach the same conclusion when implementing an even more extreme $\mstar-$stripping model in which stellar mass loss is linearly proportional to halo mass loss \citep[][Model 1]{watson_etal12}. The success of the $\rvir-$based model, together with the failure of the satellite mass loss model, suggests that the relative size of central and satellite galaxies is largely driven by the {\rm pre-infall} evolution of the satellite.

@@

>>>>

Inclusion of satellite mass loss in the model introduces correlations in the scatter: after mass loss, satellites have smaller sizes than centrals of the same $\mstar$ due to post-infall stripping.
The purple curves in Figure \ref{fig:mstarmodelclustering} show that satellite mass loss impacts $\rhalf-$dependent clustering in the expected manner. The ``small" galaxy population has a higher satellite fraction due to satellites being smaller than centrals of the same $\mstar,$ and so in this model small galaxies cluster more strongly relative to large. However, the magnitude of the effect is not sufficiently strong  to produce clustering predictions that are consistent with observations. Evidently it is difficult to strip enough mass from satellites so that the $\rhalf-$dependent clustering predictions are in agreement with observations.  We reach the same conclusion when we tested an even more extreme $\mstar-$stripping model, in which stellar mass loss is linearly proportional to halo mass loss \citep[][Model 1]{watson_etal12}. 
As we argue in \S\ref{sec:interpretation}, this supports the conclusion that the relative difference between central and satellite size is largely in place at the time of satellite infall.

@@

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/rvir_only_cen_sat_sizes.pdf}
\caption{
{\bf Relative sizes of centrals and satellites.}
In a narrow bin of halo mass $\mhalo\equiv\mpeak\approx10^{12}\msun,$ we show the distribution of model galaxy sizes for different subpopulations galaxies, as predicted by the $\rvir-$based model. The red histogram shows the sizes of satellites; the blue histogram shows host halos that have never passed inside the virial radius of a larger halo (``true centrals"); the green histogram host halos that were subhalos inside a larger at some point in their past history (``splashback centrals"). In the $\rvir-$based model, galaxy size is set by the {\em physical} size of the virial radius at the time the halo attains its peak mass, naturally resulting in smaller sizes for satellites and splashback centrals relative to true centrals of the same $\mhalo.$
}
\label{fig:censatsizehist}
\end{figure}
%-----------------------------------------------------------------------------------------------------

%-----------------------------------------------------------------------------------------
\section{A Unified Interpretation of $\rhalf-$Dependent Clustering}
\label{sec:interpretation}
%-----------------------------------------------------------------------------------------

Comparison of observed size dependence of galaxy clustering and model predictions in the previous section produced  two intriguing results. First,  model that assumes $\rhalf\propto\rvir$ predicts dependence of both the amplitude and shape of  galaxy correlation functions on size remarkably well, especially considering that the model has only one free parameter, normalization of the linear $\rhalf-\rvir$ relation. At the same time, small galaxies cluster more strongly relative to large, which seems at odds with naive expectations that within such model larger sizes correspond to larger halo masses, $\rvir\propto\mvir^{1/3}$, which should be clustered more strongly. Second, the strong dependence of galaxy clustering on size weakens to almost no dependence for samples of red and blue galaxies selected by color within a given stellar mass range. 
Below we propose a simple and unified explanation of all these results within the $\rvir$-based size model in \S\ref{subsec:censat_sizes}, and discuss the broader implications of our interpretation in \S\ref{subsec:broader_implications}.

%----------------------------------------------------------------------------
\subsection{The Relative Size of Centrals vs. Satellites}
\label{subsec:censat_sizes}
%----------------------------------------------------------------------------

Recall that in the $\rvir$-based size model, galaxy size is proportional to the virial radius of host halo at the redshift at which halo had maximum mass. Thus, we can expect that in this model satellite galaxies will have sizes smaller than sizes of central galaxies hosted by halos of similar stellar mass.  Indeed, Figure \ref{fig:censatsizehist} shows the $\rhalf$ distributions of central, satellite, and ``splashback'' central galaxies hosted by halos of the same mass $\mhalo\equiv\mpeak\approx10^{12}\msun.$ A ``splashback central"  is defined as a present-day central galaxy that used to be a satellite, i.e., its main progenitor halo passed inside the virial radius of a larger halo at some point in its past history. On the other hand, we define a ``true central" as a galaxy that has never been a satellite. The figure shows that there is a clear trend of sizes decreasing from the central to splashback and to satellite galaxies.

@@

We can also understand the observed trend of size dependence of clustering with $\mstar$ (Figure \ref{fig:clustering_ratio_upshot}) in terms of the differences of sizes of central and satellites galaxies. The satellite fraction $F_{\rm sat}(\mstar)$ decreases as $\mstar$ increases \citep[e.g.,][]{guo_etal11,reddick_etal13}. As $F_{\rm sat}$ decreases, there are fewer satellites available to preferentially weight the subsample of smaller galaxies. Thus,  high $\mstar$ samples consist mostly of central galaxies and the split by size no longer corresponds to the split into satellite- and central-dominated samples. The size dependence of clustering correspondingly weakens and nearly disappears for samples with largest $\mstar$.

Likewise, the the significantly weaker dependence of clustering on size for color-selected samples (Figure \ref{fig:colorclustering}) and its reversed sign can be naturally explained within the same framework.  For example, it is well known that satellites are redder than centrals of the same stellar mass \citep[e.g.,][]{vdB_etal08}; accordingly, red galaxies cluster much strongly than  blue galaxies \citep[e.g.,][]{zehavi_etal11}. Samples of red (blue) galaxies with stellar masses of $\mstar\sim 10^{10}\ M_\odot$ are dominated by satellite (central) galaxies and split in size within such samples no longer corresponds to the split into satellite- and central-dominated samples. In this case, $\rvir$-based size model, which reproduces the observed trend, predicts that split by size does correspond to the split by halo mass $\rhalf\propto\rvir\mhalo^{1/3}$ and that that large galaxies should be somewhat more clustered than the small ones. That is they predict that size dependence of clustering  is reversed for color-selected samples and the magnitude of the dependence is significantly weaker due to weak dependence of virial radius on halo mass and weak dependence of clustering on halo mass in this regime.

While we consider such an explanation plausible and likely, a conclusive confirmation requires proper forward modeling of size and color jointly. Such an effort is beyond the scope of, but is well-motivated by the present work.

@@

The clustering ratios in Figure \ref{fig:mstarmodelclustering} show that the impact of physically motivated levels of satellite mass loss likely cannot explain the $\rhalf-$dependent clustering seen in SDSS. The success of the $\rvir-$based model, together with the failure of the satellite mass loss model, suggests that the relative size of central and satellite galaxies is largely driven by the {\rm pre-infall} evolution of the satellite.

@@

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/rvir_only_lensing_prediction.pdf}
\caption{
{\bf Prediction for $\rhalf-$dependence of galaxy lensing.}
Using the $\rvir-$based model described in \S\ref{subsubsec:rvirmodel}, we make predictions for as-yet-unseen measurements of the $\rhalf-$dependence of galaxy lensing of $\mstar-$complete samples, where ``small" and ``large" subsamples are defined as in \S\ref{subsec:sizedef}. As a generic consequence of satellites being smaller than centrals of the same mass, we predict that future lensing measurements of $\mstar-$complete samples will show $\Delta\Sigma$ of small galaxies to be significantly stronger relative to large galaxies of the same stellar mass. To date, the $\rhalf-$dependence of $\Delta\Sigma$ has only been measured for color-selected samples \citep{charlton_etal17}, in which the trend is reversed and weakened: for both blue and red samples, $\Delta\Sigma$ of large galaxies is (slightly) stronger relative to small galaxies. 
}
\label{fig:lensingprediction}
\end{figure}
%-----------------------------------------------------------------------------------------------------
